paper_replication

### 2021-12-15 paper_replication.Rmd
Formatting tables and figures

length(unique(abortions2010$state)) #only 42 states in abortion data

#look at percent missing abortion counts by state
aggregate(X2010 ~ state, data = abortions2010, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

@@

#look at counties in abortion data but not countydem data
abortions2010$county[!(abortions2010$county %in% countydem$county)]
#"St. Louis County", "Bedford City", "Clifton Forge City"

#look at counties in countydem data but not abortion data
countydem$county[!(countydem$county %in% abortions2010$county)] #many 
unique(countydem$state.abb[!(countydem$state.abb %in% abortions2010$state)]) #"AK" "CT" "IA" "LA" "MA" "NH" "NM" "WY"

@@

#look at percent missing abortion rates by state
aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

@@

#look at percent missing dem vote by state
aggregate(percentdem2008 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

@@

#look at percent missing distance by state
aggregate(distancetoprovider ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#make new variable for high vs. low restrictions
restrictions <- restrictions %>% mutate(totalrestrict = parental_involvement + two_trips + trap + unconstitutional_ban, 
                                        restrictionlevel = case_when(totalrestrict >= 3 ~ "high", 
                                                              totalrestrict < 3 ~ "low"))
restrictions$state[!(restrictions$state %in% countydem$state)] #none
countydem <- left_join(countydem, restrictions, by = "state")
countydem$restrictionlevel <- factor(countydem$restrictionlevel, levels = c("low", "high"))

@@

summarytable <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydemanalysis, function(x) {sum(!(is.na(x)))}, na.action = NULL)
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Abortion Data Included in Analyses by State") %>% kable_styling()
# names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | inanalysislabel, 
       data = countydem,
       extra.col = list(`P-value` = p_value_t_test),
       footnote = 'P-values are derived by two sample t-tests.'
       )
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#calculate propensity scores
propensitywts <- glm(inanalysis ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008, data=countydem, family=binomial())
summary(propensitywts)

@@

#3 plots of distribution of propensity scores

ggplot(df, aes(x = ps, group = inanalysislabel)) + geom_density(aes(fill = inanalysislabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Figure 1: Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")

@@

model_dist <- lm(AbortionRate2010 ~ restrictionlevel + distancetoprovider + restrictionlevel*distancetoprovider, data = countydemanalysis, weights = ps)
summ_dist <- summary(model_dist)

# Model Summaries (Note coefficients are per 1000 women)
table_3 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2]*1000,
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "Table 3: Change in Abortions per 1000 Women in Model without Distances", 
          digits = 3)
print(table_3)

p <- summ_dist$coefficients[4,4] %>% round(3)
table_4 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2]*1000, model_dist$coefficients[3]*1000,
                           model_dist$coefficients[4]*1000),
                  "p-value" = c('< 0.001', '< 0.001', p)) %>%
  kbl(caption = "Table 4: Change in Abortions per 1000 Women in Model with Distances")
print(table_4)

@@

#calculate propensity scores
#include education this time
propensitywts <- glm(hasabortiondata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
summary(propensitywts)

@@

#3 plots of distribution of propensity scores

ggplot(df, aes(x = ps, group = hasabortiondatalabel)) + geom_density(aes(fill = hasabortiondatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")

ggplot(df, aes(x = ps, fill = hasabortiondatalabel, color = hasabortiondatalabel)) +
  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
  facet_grid(as.factor(hasabortiondata) ~ .) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  scale_color_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')

df %>%
  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
  group_by(hasabortiondata, ps.grp) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(n2 = ifelse(hasabortiondata == 0, yes = n, no =  -1*n)) %>%
  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasabortiondata))) +
  geom_bar(stat = 'identity', position = 'identity') +
  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasabortiondata == 0, 8, -8))) +
  xlab('Probability of Being in the Analysis') +
  ylab('N') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete(name = "", labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data")) +
  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
  theme(legend.position = 'bottom', legend.direction = 'vertical',
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

@@

# Model Summaries (Note coefficients are per 1000 women)
table_ext_no_dist <- tibble(Characteristic = "Number of Types of Restrictive Legislation",
                  Beta = ext_model_no_dist$coefficients[2]*1000,
                  "p-value" = summ_ext_no_dist$coefficients[2,4]) %>% 
  kbl(caption = "Table ****: Change in Abortions per 1000 Women in Model without Distances, all 50 States",
          digits = 3)
print(table_ext_no_dist)

table_ext_dist <- tibble(Characteristic = c("Number of Types of Restrictive Legislation", 
                                            "Distance to Provider", 
                                            "Distance/Restriction Interaction"),
                  Beta = ext_model_dist$coefficients[2:4]*1000,
                  "p-value" = summ_ext_dist$coefficients[2:4,4]) %>% 
  kbl(caption = "Table ****: Change in Abortions per 1000 Women in Model with Distances, all 50 States", digits = 3)
print(table_ext_dist)

@@

print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasinfantdeathdatalabel, 
       data = countydem,
       extra.col = list(`P-value` = p_value_t_test),
       footnote = 'P-values are derived by two sample t-tests.'
       ))

#calculate propensity scores
#include education this time
propensitywts <- glm(hasinfantdeathdata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())

@@

#3 plots of distribution of propensity scores

plot1 <- ggplot(df, aes(x = ps, group = hasinfantdeathdatalabel)) + geom_density(aes(fill = hasinfantdeathdatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")
print(plot1)

@@

model <- lm(infantdeathrate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydeminfantdeathanalysis, weights = ps)
summary(model)

@@

#calculate propensity scores
#include education this time
propensitywts <- glm(hasbirthratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
summary(propensitywts)

p.score <- predict(propensitywts, countydem, type = "response")
df <- countydem
df$ps <- p.score
#3 plots of distribution of propensity scores

plot1 <- ggplot(df, aes(x = ps, group = hasbirthratedatalabel)) + geom_density(aes(fill = hasbirthratedatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")
print(plot1)

@@

model <- lm(Birth.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydembirthrateanalysis, weights = ps)
summary(model)

@@

#calculate propensity scores
#include education this time
propensitywts <- glm(hasfertilityratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
summary(propensitywts)

p.score <- predict(propensitywts, countydem, type = "response")
df <- countydem
df$ps <- p.score
#3 plots of distribution of propensity scores

@@

model <- lm(Fertility.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemfertilityrateanalysis, weights = ps)
summary(model)

>>>>

# length(unique(abortions2010$state)) #only 42 states in abortion data

#look at percent missing abortion counts by state
# aggregate(X2010 ~ state, data = abortions2010, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

@@

#look at counties in abortion data but not countydem data
# abortions2010$county[!(abortions2010$county %in% countydem$county)]
#"St. Louis County", "Bedford City", "Clifton Forge City"

#look at counties in countydem data but not abortion data
# countydem$county[!(countydem$county %in% abortions2010$county)] #many 
# unique(countydem$state.abb[!(countydem$state.abb %in% abortions2010$state)]) #"AK" "CT" "IA" "LA" "MA" "NH" "NM" "WY"

@@

#look at percent missing abortion rates by state
# aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

@@

#look at percent missing dem vote by state
# aggregate(percentdem2008 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

@@

#look at percent missing distance by state
# aggregate(distancetoprovider ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#make new variable for high vs. low restrictions
restrictions <- restrictions %>% mutate(totalrestrict = parental_involvement + two_trips + trap + unconstitutional_ban, 
                                        restrictionlevel = case_when(totalrestrict >= 3 ~ "high", 
                                                              totalrestrict < 3 ~ "low"))
# restrictions$state[!(restrictions$state %in% countydem$state)] #none
countydem <- left_join(countydem, restrictions, by = "state")
countydem$restrictionlevel <- factor(countydem$restrictionlevel, levels = c("low", "high"))

@@

table_1 <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydemanalysis, function(x) {sum(!(is.na(x)))}, na.action = NULL)
table_1 <- left_join(table_1, usedcounties, by = "state")
table_1 %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Number of Counties with Abortion Data Included in Analyses by State", digits = 3) %>% kable_styling()

to_adjust_kable_table_numbers <- kable(table_1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | inanalysislabel, 
       data = countydem,
       extra.col = list(`P-value` = p_value_t_test),
       footnote = 'P-values are derived by two sample t-tests.',
       caption = "County Data for States Included and Not Included in Analysis"
       )
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#calculate propensity scores
propensitywts <- glm(inanalysis ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008, data=countydem, family=binomial())
# summary(propensitywts)

@@

# plot of distribution of propensity scores 
ggplot(df, aes(x = ps, group = inanalysislabel)) + geom_density(aes(fill = inanalysislabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Figure 1: Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")

@@

model_dist <- lm(AbortionRate2010 ~ restrictionlevel + distancetoprovider + restrictionlevel*distancetoprovider, data = countydemanalysis, weights = ps)
summ_dist <- summary(model_dist)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Model Summaries (Note coefficients are per 1000 women)
table_3 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2]*1000,
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "Change in Abortions per 1000 Women in Model without Distances", 
          digits = 3)
table_3
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p <- summ_dist$coefficients[4,4] %>% round(3)
table_4 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2]*1000, model_dist$coefficients[3]*1000,
                           model_dist$coefficients[4]*1000),
                  "p-value" = c('< 0.001', '< 0.001', p)) %>%
  kbl(caption = "Change in Abortions per 1000 Women in Model with Distances")
table_4

@@

#calculate propensity scores
#include education this time
propensitywts <- glm(hasabortiondata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
# summary(propensitywts)

@@

# plot of distribution of propensity scores 
ggplot(df, aes(x = ps, group = hasabortiondatalabel)) + geom_density(aes(fill = hasabortiondatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")

@@

# Model Summaries (Note coefficients are per 1000 women)
table_ext_no_dist <- tibble(Characteristic = "Number of Types of Restrictive Legislation",
                  Beta = ext_model_no_dist$coefficients[2]*1000,
                  "p-value" = summ_ext_no_dist$coefficients[2,4]) %>% 
  kbl(caption = "Change in Abortions per 1000 Women in Model without Distances, all 50 States", digits = 3)
table_ext_no_dist
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_ext_dist <- tibble(Characteristic = c("Number of Types of Restrictive Legislation", 
                                            "Distance to Provider", 
                                            "Distance/Restriction Interaction"),
                  Beta = ext_model_dist$coefficients[2:4]*1000,
                  "p-value" = summ_ext_dist$coefficients[2:4,4]) %>% 
  kbl(caption = "Change in Abortions per 1000 Women in Model with Distances, all 50 States", digits = 3)
table_ext_dist
```

@@

print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasinfantdeathdatalabel, 
       data = countydem,
       extra.col = list(`P-value` = p_value_t_test),
       footnote = 'P-values are derived by two sample t-tests.'
       ))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#calculate propensity scores
#include education this time
propensitywts <- glm(hasinfantdeathdata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())

@@

# plot of distribution of propensity scores 
plot1 <- ggplot(df, aes(x = ps, group = hasinfantdeathdatalabel)) + geom_density(aes(fill = hasinfantdeathdatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")
plot1

@@

model <- lm(infantdeathrate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydeminfantdeathanalysis, weights = ps)
# summary(model)

@@

#calculate propensity scores
#include education this time
propensitywts <- glm(hasbirthratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
# summary(propensitywts)

# plot of distribution of propensity scores

plot1 <- ggplot(df, aes(x = ps, group = hasbirthratedatalabel)) + geom_density(aes(fill = hasbirthratedatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")
plot1

@@

model <- lm(Birth.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydembirthrateanalysis, weights = ps)
# summary(model)

@@

#calculate propensity scores
#include education this time
propensitywts <- glm(hasfertilityratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
# summary(propensitywts)

p.score <- predict(propensitywts, countydem, type = "response")
df <- countydem
df$ps <- p.score
# plot of distribution of propensity scores

@@

model <- lm(Fertility.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemfertilityrateanalysis, weights = ps)
# summary(model)
